---
title: "Betke-Briana-ADA-homework-1"
author: "Briana Betke"
date: "2/9/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 1
This is an Rmarkdown of my work for challenges 1-4 of homework 1! YAY!

#### Step 1. Create a vector called quote. Remove punctuation because I am *ambitious*. 
```{r challenge 1}
# Create quote vector
quote <- c("There is grandeur in this view of life, with its several powers, having been originally breathed by the Creator into a few forms or into one; and that, whilst this planet has gone circling on according to the fixed law of gravity, from so simple a beginning endless forms most beautiful and most wonderful have been, and are being evolved.")
# Removing punctuation
quote <- gsub("[[:punct:]]","", quote)
quote
```

#### Step 2. Split the vector into a new vector called split_quote.
```{r}
library(stringr)
# Split the vector by word.
split_quote <- str_split(quote, " ")
split_quote
```

#### Step 3. Extract every 4th element from the Vector. Name it every_fourth. 
```{r}
# Take every forth word
every_fourth <- split_quote[[1]][seq(from = 1, to = 60, by = 4)]
every_fourth
```

#### Step 4. Rearrange every_forth in reverse alphabetical order.
```{r}
# Using the sort function, make sure to indicate TRUE on decreasing.
sort(every_fourth, decreasing = TRUE)
```


### Challenge 2: Adventures with Matracies
#### Step 1. Create Matrix that contains 8 rows and 10 columns, named m1. Must have odd integers from 1 to 159.
```{r}
# Using a sequence to fill matrix with odd integers.
m1 <- matrix(data = seq(from = 1, to = 159, by = 2), nrow = 8, ncol = 10)
m1
```

#### Step 2. Extract and print the elements in row 5, column 2 with *single bracket notation*.
```{r}
m1[5,2] # Indicate row 5 and column 2 in single brackets
```

#### Step 3. Extract and print the contents of rows 5 to 7 of the matrix.
```{r}
m1[5:7, ]
```

#### Step 4. Extract the contents of (row 3, column 4) to (row 6, column 9) as a new variable, m2. Find the class or mode.
```{r}
# Extract the rows and columns, name it m2
m2 <- m1[3:6,4:9]
m2 # View the matrix

# Get the class of m2
class(m2)

# Get the mode
mode(m2)
```

### Challenge 3: No Dissarry with this Array
#### Step 1. Create an array with 4 dimensions and 400 elements with numbers 400 to 1.
```{r}
# Create array. Place teh 400 before 1 for the data to be in decreasing order.
a <- array(data = 400:1, dim = c(5, 5, 4, 4))
a
```

#### Step 2. What would return with the following code?
```{r}
a[1, 1, 1, 2]

a[2, 3, 2, ]

a[1:5, 1:5, 3, 3]
```

### Challenge 4: Torture with Lists, but Primates are Cute. So it makes it ok.
To create the list of primates, I started by making character vectors of the families. Then those families were stored to respective infraorders. Then the respective infraorders were combined into two lists, representing the two suborders Strepsirhini and Haplorini. Those were then stored together as primates, the order. ~~I tried!~~

#### Step 1. Creating Strepsirhini

Starting with the infraorder lorisiformes
```{r}
# Make a list for the infraorder lorisiformes 
lorisFam <- c("Lorisidae", "Galagidae") # create a character vector of superfamily 
lorisSuper <- list(lorisFam) # Make a list
names(lorisSuper) <- "lorisoidea" # Name it super family name
lorisSuper # View the list 
```

Then infraorder lemuriformes
```{r}
# Do the same for lemuriformes
lemurFam<- c("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae")
lemurSuper <- list(lemurFam)
names(lemurSuper) <- "Lemuroidea"
lemurSuper 
```

Put both infraorders together to make suborder Strepsirhini
```{r}
# Suborder strepsirhini 
strepsirhini <- list(lorisSuper,lemurSuper) # make a list of superfamilies
names(strepsirhini) <- c("lorisiformes","lemuriformes") # Name them by infraorder
strepsirhini
```

#### Step 2. Creating Haplorhini

Starting with Tarsiiformes
```{r}
# Tarsiiformes
tarsirFam <- "Tarsiidae"
tarsirSuper <- list(tarsirFam)
names(tarsirSuper) <- "Tarsioidea"
tarsirSuper
```

Then the Simiiformes. 
Beginning with platyrrhini parvorder.
```{r}
# Parvorder: Platyrrhini
ceboFam <- c("Cebidae", "Atelidae", "Pitheciidae")
ceboSuper <- list(ceboFam)
names(ceboSuper) <- "Ceboidea"
ceboSuper
```

Catarrhini Next.
```{r}
# Parvorder catarrhini
# superfam Hominids
hominFam <- c("Hylobatidae", "Hominidae")
hominSuper <- list(hominFam)
names(hominSuper) <- "Hominoidea"
hominSuper

# other superfam
cercFam <- "Cercopithecidae"
cercSuper <- list(cercFam)
names(cercSuper) <- "Cercopithecoidea"
cercSuper
```

Putting parvorders together to make Catarrhini
```{r}
# create list for Catarrhini
catarr <- list(hominSuper, cercSuper)
catarr
```

Put the infraorders together
```{r}
# Create list for Simiiformes
simiInfra <- list(ceboSuper,catarr)
names(simiInfra) <- c("Platyrrhini","Catarrhini")
simiInfra
```

The final step to make Haplorhini
```{r}
# combine Tarsiiformes and Simiiformes
haplor <- list(tarsirSuper,simiInfra)
names(haplor) <- c("Tarsiiformes","Simiiformes")
haplor
```

#### Step 3. Combining the suborders and naming it primate
```{r}
# put Strepsirhini and Haplorhini together and call it primates
primates <- list(strepsirhini,haplor)
names(primates) <- c("Strepsirhini","Haplorhini")
primates
```

#### Step 4. Extract Platyrrhini and name it nwm.
```{r}
# Extract Platyrrhini and name it nwm.
nwm <- primates[["Haplorhini"]][["Simiiformes"]][["Platyrrhini"]]
nwm
```

#### Step 5.Class and mode of nwm.

First, get the class.
```{r}
# Get the class of nwm
class(nwm)
```

Then the mode.
```{r}
# Get the Mode of nwm
mode(nwm)
```

#### Step 6. $ sign notation to extract tarsirs.
```{r}
# $ sign notation to extract tarsirs
primates$Haplorhini$Tarsiiformes$Tarsioidea
```

